### 2019-09-23 Manuscript/Manuscript.tex
Updates from Overleaf

Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo values of 598--729~ppm (5--95 percentile), and the ``reduced I\&NG + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig. \ref{fig:scenario_pCO2}). These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm. Estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between proxy records and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a}. These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene. The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}. These modeled values are therefore consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.

>>>>

Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo values of 598--729~ppm (5--95 percentile), and the ``reduced I\&NG + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig. \ref{fig:scenario_pCO2}). These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm. Estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between proxy records and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}. These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene. The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}. These modeled values are therefore consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.

### 2019-09-23 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
    \caption{Modeled vs. data-constrained \COtwo consumption in watersheds around the world. \textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day. Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations. The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines. \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.}
    \label{fig:r2_cross_plot}
\end{figure}

@@

The decrease in \pCOtwo that we estimate associated with the emergence of I\&NG since 5~Ma ($\sim$120~ppm) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm). This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}). \pCOtwo is then estimated from $T$ using a simple energy balance equation. However, as discussed in the main text, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo. The linear relationship described in Equation \ref{eq:Molnar3} ignores these feedbacks in the global climate system.

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
    \caption{Modeled vs. data-constrained \COtwo consumption in watersheds around the world. \textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day. Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations. The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines. \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$. The chemical weathering and regolith thickness maps of this parameter combination are shown in Figure \ref{fig:weathering_regolith_maps}.}
    \label{fig:r2_cross_plot}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
    \caption{Chemical weathering and regolith thickness maps for the parameter combination that produced that highest $r^{2}$ (Fig. \ref{fig:r2_cross_plot}B). Parameter values used for this combination are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
    \label{fig:weathering_regolith_maps}
\end{figure}

@@

@@

The decrease in \pCOtwo that we estimate associated with the emergence of I\&NG since 5~Ma ($\sim$120~ppm) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm). This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}). \pCOtwo is then estimated from $T$ using a simple energy balance equation. However, as discussed in the main text, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo. The linear relationship used in \citet{Molnar2015a} ignores these feedbacks in the global climate system.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/pCO2_proxies.pdf}
    \caption{Proxy-based \pCOtwo estimates for the past 20~m.y. For the \citet{Foster2017a} data, error bars indicate standardized uncertainties. For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates. The \citet{Bereiter2015a} data comes from the EPICA Dome C ice core.}
    \label{fig:pCO2_proxies}
\end{figure}

Proxy-based \pCOtwo estimates for the past 20~m.y. are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a}. Even when stringent quality criteria and the latest understanding of each of the \pCOtwo proxies have been applied to available \pCOtwo records, both significant uncertainty in the estimated \pCOtwo for any given data point as well as disagreement between techniques remain. Nevertheless, our modeled \pCOtwo values for 15~Ma (Fig. \ref{fig:scenario_pCO2_all}) resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene.
